---
title: "MWF_HW1"
author: "Savannah McNair"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

necessary packages
```{r}
require(dplyr)
require(tidyr)
```


# 1. Set-up of Git and GitHub (20%)

install Git on your computer
- I already had Git downloaded to my computer! Check!

make GitHub account (get the free pro version with an educational account)
- I already have a GitHub account - @savmcnair . Check!

set-up Git in Rstudio (if you get stuck you can try using GitHub Desktop)
- I already have Git setup with RStudio, as I use Git and R for work. I noted this in the discussion forum, but I use GitHub desktop to sync my local computer, R, and GitHub repos. Check!

#2. Do online GitHub course (20%)
https://github.com/skills/introduction-to-github

- Created a the skills intro repo and cloned to my local comp (link to repo here: https://github.com/savmcnair/skills-introduction-to-github/tree/main)
- Created a new branch, sav-newbranch
- Added a PDF to the repo, push
- Created pull request with new commit
- Merged pull request (main and sav-newbranch)
- Check!

#3. Do your first Git project (total of 60%)
create a new project and set-up the project structure (folder, README, etc.) (5%)
- Created a new repo (link to repo here: https://github.com/savmcnair/RStudio-MWF-repo)
- Has a README and is in a folder linked to my desktop

download two dataset about COVID-19 from this GitHub repository: “UID_ISO_FIPS_LookUp_Table”
and “time_series_covid19_confirmed_global”. Save both data locally (5%)
- downloaded both from the repository and added them to my own repository, above
- committed this change 

merge the two datasets and create a long version of the data. Save both long and wide data locally
(20%)

read in the datasets
```{r}
UIDs <- read.csv("C:/Users/smcnair1/Desktop/RStudio-MWF-repo/UID_ISO_FIPS_LookUp_Table.csv")
global_covid <- read.csv("C:/Users/smcnair1/Desktop/RStudio-MWF-repo/time_series_covid19_confirmed_global.csv")
```

merge the datasets based on lat
```{r}
merged_df <- merge(x = UIDs, y = global_covid, by = "Lat", all = TRUE)
```

make long df
```{r}
merged_df_longer <- merged_df %>%
  pivot_longer(
    cols = starts_with("X"), 
    names_to = "XTime",
    values_to = "XVal")
```

save long and wide locally
```{r}
save(merged_df, file = "C:/Users/smcnair1/Desktop/RStudio-MWF-repo/wide_data.RData")
save(merged_df_longer, file = "C:/Users/smcnair1/Desktop/RStudio-MWF-repo/long_data.RData")
```


create three graphs using ggplot and save them (10%):
i) overall change in time of log number of cases
ii) change in time of log number of cases by country
iii) change in time by country of rate of infection per 100,000 cases

use the README file to write a mini-report (10%). 
It should include:
a. description of the project (what you aim to do, where you got the data from, etc.)
b. explain the organization of the repo (folder structure, where they can find the data and scripts,
what were the steps in creating the report)
c. main findings where you include the three graphs and a sentence or two on their interpretation
d. session info to help with reproducibility

presentation (10%). Clear structure of the folders, at least three commits with appropriate comments,
well formatted README and graphs

Keep you repositories (for tasks 2 and 3) private. For your grading you will invite the teacher
(@alex-cernat) to be collaborators to your repository. In that way we can mark your work.
1
